---
title: "Factor Analysis"
author: "Rajitha Senanayake"
date: "7/26/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = FALSE)
```

## Load Libraries
```{r}
library(corrplot)
library(psych)
library(lavaan)
library(factoextra)
```

## Load Dataset
```{r}
data <- read.csv('C:\\Users\\Rajiy\\Downloads\\customers.csv')
head(data)
```

## Cleaning Data
```{r}
sales <- data[,3:8]
head(sales)
```
## Correlation Matrix and Heatmap Plot
```{r}
cor(sales, method = 'spearman')
```
```{r}
corrplot(cor(sales, method = 'spearman'), method = 'number')
```
## Standardizing Variables
```{r}
scaled_df <- scale(sales, center = TRUE, scale = TRUE)
head(scaled_df)
```
## Bartlett Test to Check usefulness of Factor Analysis
```{r}
cortest.bartlett(scaled_df)
```

## Calculate the Eigen Values and Eigen Vectors
```{r}
sales.cov <- cov(scaled_df)
sales.eigen <- eigen(sales.cov)
sales.eigen
```
*  Number of eigen values > 1
```{r}
x <- sales.eigen$values
sum(x > 1)
```
* The number of factors to be considered m=2, we can confirm this using a scree plot
```{r}
plot(sales.eigen$values, xlab = 'Eigenvalue Number', ylab = 'Eigenvalue Size', main = 'Scree Graph', type = 'b', xaxt = 'n')
axis(1, at = seq(1, 6, by = 1))
```
## Resulted Factors
```{r}
sales.factors <- sales.eigen$vectors[,1:2]
sales.factors <- -sales.factors
row.names(sales.factors) <- c('Fresh','Milk','Grocery','Frozen','Detergents_Paper','Delicatessen')
colnames(sales.factors) <- c('Factor_1','Factor_2')
sales.factors
```
## Loading Matrix
```{r}
C <- as.matrix(-sales.eigen$vectors[,1:2])
D <- matrix(0, dim(C)[2], dim(C)[2])
diag(D) <- sales.eigen$values[1:2]
sales.loadings <- C %*% sqrt(D)
row.names(sales.loadings) <- c('Fresh','Milk','Grocery','Frozen','Detergents_Paper','Delicatessen')
colnames(sales.loadings) <- c('Fac_1_loadings','Fac_2_loadings')
sales.loadings
```

## Calculating Communalities
```{r}
h2 <- rowSums(sales.loadings^2)
h2
```


* Total communality
```{r}
sum(h2)
```
## Proportion of Total Variance Explained by the Two Factors
```{r}
sum(h2)/6
```

## Specific Variance Matrix
```{r}
specVar <- diag(cov(scaled_df))-h2
specVar
```

## Varimax Factor Rotation
```{r}
varimax(sales.loadings)
```

## Test for sufficient factors
```{r}
fa2 <- fa(scaled_df, nfactors = 2, rotate = 'none')
factor.stats(scaled_df, fa2)
```
## Confirmatory Factor Analysis with Lavaan
```{r}
HS.model <- ' essentials =~  Milk + Grocery + Detergents_Paper 
              storedtemp =~ Fresh + Frozen + Delicatessen'
```

```{r}
fit <- cfa(HS.model, data=scaled_df)
fit
```
```{r}
summary(fit, fit.measures=TRUE)
```
```{r}
cortest.bartlett(scaled_df)
```
```{r}

```

